---
title: "HW2"
author: "Ekta Chaudhary"
date: "20/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)
```

```{r}
library(tidyverse)
library(caret)
library(ModelMetrics)
library(glmnet)
library(gam)
library(boot)
library(mgcv)
library(splines)
library(ggplot2)
library(lasso2)
```

Reading the Datasets 

```{r}
data = 
  read_csv('./data/College.csv') %>%
select(-College)
train =
  data[-125,]
test =
  data[125,]
```

```{r}
x <- model.matrix(Outstate~.,train)[,-1]
y <- train$Outstate
```

(a) Create scatter plots of response vs. predictors.

```{r, fig.height=4}
theme1 <- trellis.par.get()
theme1$plot.symbol$col <- rgb(.3, .5, .2, .5)
theme1$plot.symbol$pch <- 18
theme1$plot.line$col <- rgb(.8, .1, .1, 1)
theme1$plot.line$lwd <- 2
theme1$strip.background$col <- rgb(.0, .2, .6, .2)
trellis.par.set(theme1)
featurePlot(x, y, plot = "scatter", labels = c("","Y"),
            type = c("p"), layout = c(4, 2))
```

We will now fit univariate nonlinear models and GAM to investage the association between `Outstate` and `Terminal`.
The function `poly()` returns a matrix whose columns are a basis of orthogonal polynomials, which essentially means that each column is a linear combination of `Terminal`, `Terminal^2`, `Terminal^3`, and `Terminal^4`.

```{r}
fit1 <- lm(Outstate~Terminal, data = train)
fit2 <- lm(Outstate~poly(Terminal,2), data = train) 
fit3 <- lm(Outstate~poly(Terminal,3), data = train)
fit4 <- lm(Outstate~poly(Terminal,4), data = train) 
fit5 <- lm(Outstate~poly(Terminal,5), data = train)
```

Using `anova()` to test the null hypothesis that a simpler model is sufficient to explain the data against the alternative hypothesis that a more complex model is required. In order to use ANOVA, the models must be nested.

```{r}
anova(fit1,fit2,fit3,fit4,fit5)
```

b) Fit a smoothing spline model using Terminal as the only predictor of Outstate for a range of degrees of freedom, as well as the degree of freedom obtained by generalized cross- validation, and plot the resulting fits. Describe the results obtained.


```{r}

Terminallims <- range(train$Terminal)
Terminal.grid <- seq(from = Terminallims[1],to = Terminallims[2])

fit.ss <- smooth.spline(train$Terminal, train$Outstate)
fit.ss$df

pred.ss <- predict(fit.ss,
                   x = Terminal.grid)

pred.ss.df <- data.frame(pred = pred.ss$y,
                         Terminal = Terminal.grid)

p <- ggplot(data = train, aes(x = Terminal, y = Outstate)) +
  geom_point(color = rgb(.2, .4, .2, .5))
p + geom_line(aes(x = Terminal, y = pred), data = pred.ss.df, 
              color = rgb(.8, .1, .1, 1)) + theme_bw() + 
  labs(title = paste(round(fit.ss$df), 'degrees of freedom from GCV')) + 
  theme(plot.title = element_text(hjust = 0.5))
```