---
title: "HW2"
author: "Ekta Chaudhary"
date: "20/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)
```

```{r}
library(tidyverse)
library(caret)
library(ModelMetrics)
library(glmnet)
library(gam)
library(boot)
library(mgcv)
library(splines)
library(ggplot2)
library(lasso2)
library(pdp)
library(earth)
```

Reading the Datasets 

```{r}
data = 
  read_csv('./data/College.csv') %>%
select(-College)
data_1 =
  data[-125,]
data_2 =
  data[125,]
```

```{r}
x <- model.matrix(Outstate~.,data_1)[,-1]
y <- data_1$Outstate
```

# (a) Create scatter plots of response vs. predictors.

```{r, fig.height=4}
theme1 <- trellis.par.get()
theme1$plot.symbol$col <- rgb(.3, .5, .2, .5)
theme1$plot.symbol$pch <- 18
theme1$plot.line$col <- rgb(.8, .1, .1, 1)
theme1$plot.line$lwd <- 2
theme1$strip.background$col <- rgb(.0, .2, .6, .2)
trellis.par.set(theme1)
featurePlot(x, y, plot = "scatter", labels = c("","Y"),
            type = c("p"), layout = c(4, 2))
```

# b) Fit a smoothing spline model using Terminal as the only predictor of Outstate for a range of degrees of freedom, as well as the degree of freedom obtained by generalized cross- validation, and plot the resulting fits. Describe the results obtained.


```{r}

Terminallims <- range(data_1$Terminal)
Terminal.grid <- seq(from = Terminallims[1],to = Terminallims[2])

fit.ss <- smooth.spline(data_1$Terminal, data_1$Outstate)
fit.ss$df

pred.ss <- predict(fit.ss,
                   x = Terminal.grid)

pred.ss.df <- data.frame(pred = pred.ss$y,
                         Terminal = Terminal.grid)

p <- ggplot(data = data_1, aes(x = Terminal, y = Outstate)) +
  geom_point(color = rgb(.2, .4, .2, .5))
p + geom_line(aes(x = Terminal, y = pred), data = pred.ss.df, 
              color = rgb(.8, .1, .1, 1)) + theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5))
```

The function `smooth.spline()` can be used to fit smoothing spline models. Generalized cross-validation is used to select the degree of freedom. The degree of freedom obtained by generalized cross- validation is **4.468629**. 

```{r}
par(mfrow = c(3,3)) # 3 x 3 grid
all.dfs = rep(NA, 9)
for (i in 2:20) {
  fit.ss = smooth.spline(data_1$Terminal, data_1$Outstate, df = i)
  
  pred.ss <- predict(fit.ss, x = Terminal.grid)
  
  plot(data_1$Terminal, data_1$Outstate, cex = .5, col = "darkgrey")
  title(paste("Degrees of freedom = ", round(fit.ss$df)),  outer = F)
  lines(Terminal.grid, pred.ss$y, lwd = 2, col = "blue")
}
```
I have picked a range of degrees of freedom from 2 to 20. As it can be seen from the plots, when the degree of freedom is 2, the model is linear and when the df increases the model gets wiggly.

# c) Fit a generalized additive model (GAM) using all the predictors. Plot the results and explain your findings.

```{r}
gam.m1 = gam(
  Outstate~ Apps + Accept + Enroll + Top10perc + Top25perc + F.Undergrad + P.Undergrad + Room.Board + Books + Personal + PhD + Terminal + S.F.Ratio + perc.alumni + Expend + Grad.Rate, 
  data = data_1)
gam.m2 = gam(
  Outstate~ Apps + Accept + Enroll + Top10perc + Top25perc + F.Undergrad + P.Undergrad + Room.Board + Books + Personal + PhD + s(Terminal) + S.F.Ratio + perc.alumni + Expend + Grad.Rate,
  data = data_1)
gam.m3 = gam(
  Outstate~ Apps + Accept + Enroll + Top10perc + Top25perc + F.Undergrad + P.Undergrad + te(Room.Board) + te(Personal) + Books + PhD + s(Terminal) + S.F.Ratio + perc.alumni + Expend + Grad.Rate, data = data_1)

anova(gam.m1, gam.m2, gam.m3, test = "F")

```

Looking at the p-values from the ANOVA test, Model 3 appears to be the best fitting model.

```{r}
plot(gam.m2)
plot(gam.m3)
vis.gam(gam.m3, view = c("Room.Board","Personal"),plot.type = "contour", color = "topo")
```

(d) Fit a multivariate adaptive regression spline (MARS) model using all the predictors. Report the final model. Present the partial dependence plot of an arbitrary predictor in your final model.

```{r}
ctrl1 <- trainControl(method = "cv", number = 10)
mars_grid <- expand.grid(degree = 1:2, 
                         nprune = 2:10)

set.seed(2)
mars.fit <- train(x, y,
                  method = "earth",
                  tuneGrid = mars_grid,
                  trControl = ctrl1)

ggplot(mars.fit)

mars.fit$bestTune

coef(mars.fit$finalModel) 
```

Present the partial dependence plot of an arbitrary predictor in your final model.

```{r}
p1 <- partial(mars.fit, pred.var = c("Room.Board"), grid.resolution = 10) %>% autoplot()

p2 <- partial(mars.fit, pred.var = c("Room.Board", "perc.alumni"), grid.resolution = 10) %>%
      plotPartial(levelplot = FALSE, zlab = "yhat", drape = TRUE, 
                  screen = list(z = 20, x = -60))

grid.arrange(p1, p2, ncol = 2)
```


```{r}
pred.gam <- predict(gam.m2, newdata = data_2)
pred.mars <- predict(mars.fit, newdata = data_2)
pred.gam
pred.mars
```

